Reinforcement Learning for Options on Target Volatility Funds

### 2021-05-19 tex/targetvol.tex
Paper

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
%\usepackage{marginnote}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx}
\usepackage{graphicx,tikz}
\usepackage{draftwatermark}
\usepackage{enumerate} 
% \usepackage{listings}
% \usepackage{booktabs}
\usepackage{mathtools}
% \usepackage[gen]{eurosym}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}

% \setbeamertemplate{bibliography item}[text]
% \renewcommand*{\bibfont}{\scriptsize}

@@

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Milan, roberto.daluiso@intesasanpaolo.com}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A., emanuele.nastasi@marketz.eu}
	\and Andrea Pallavicini\thanks{Imperial College London and Intesa Sanpaolo Milan, a.pallavicini@imperial.ac.uk} 
	\and
	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com}% indicates same as 2nd thanks
}

@@

In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard28} and \textcite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim}. Moreover, in pricing literature derivative contracts on TVS are studied, see in particular \textcite{Di_Graziano}, \textcite{Grasselli}, and \textcite{Albeverio}.

In this contribution we study a problem related to TVS. We deal with the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a protection to a portfolio manager on the capital invested in a TVS. the portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model. Then, we tackle the problem in the general case by using reinforcement learning (RL) algorithms.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we derive the analytical results, and in \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work.

@@

    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}},

@@

    \ell \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\|\alpha_t \cdot \nu_t \|}\bigg|I_t = K\right],

@@

\subsection{Black and Scholes Model}
In the \textcite{BlackScholes} model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift can be written as
\begin{equation}
     \ell \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\|\alpha\left(t,K\right) \cdot \nu\left(t\right)\|},  
\end{equation}
so that the optimization problem can be solved looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic function of time.

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, indipendently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}\end{equation}
\end{proposition}
The absence of stochastic elments in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori}; in fact one can solve the optimization problema once for all for each $t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{BS} = BS(F^{\text{TVS}}(0,T;\alpha^*), K, T, \bar{\sigma}, D(0,T;\zeta)),
\end{equation}
where $F^{\text{TVS}}(t,T;\alpha)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha) = I_t \exp\left[\int_t^T\left(r(u)-\bar{\sigma}\frac{\alpha(u)\cdot \mu(u)}{\|\alpha(u)\cdot \nu(u)\|}\right)du\right],
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Black's Closed form solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \nu^T$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

\begin{lemma}
	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \nu^{T}$. Then
	\begin{equation}\label{eq:inf}
		\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\alpha \cdot \mu}{\norm{\alpha \cdot \nu}} = \min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}};
	\end{equation}
	if $\bar{\imath}$ is the index which realizes the min, then the infimimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
\end{lemma}
\begin{proof}
	Let us first consider the case in which $\mu = \mathbbm{1}$. Since the argument of the infimum is zero-homogeneous, normalizing by $\alpha \cdot \mathbbm{1} > 0$ we can restrict to the affine hyperspace $\{\alpha \cdot \mathbbm{1} = 1\}$, where the minimization \eqref{eq:inf} reduces to the maximization of its denominator: the required infimum will be square root of the reciprocal of
	\[
	\sup\left\{\norm{\alpha \cdot \nu}^2 \mid \alpha \in \mathbb{R}_{+}^n,\, \alpha \cdot \mathbbm{1} = 1\right\}.
	\]
	Now we can note that $\Sigma$ is positive definite, hence $\Sigma_{ij} < \sqrt{\Sigma_{ii}\Sigma_{jj}} \leq \Sigma_{\bar{\imath}\bar{\imath}}$, which implies
	\[
	\norm{\alpha \cdot \nu}^2 = \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{ij} \leq \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{\bar{\imath}\bar{\imath}} = \Sigma_{\bar{\imath}\bar{\imath}}
	\]
	because $\sum_i \alpha_i = 1$. Since we trivially have equality for $\alpha_i = \delta_{i\bar{\imath}}$, this concludes the proof of the case $\mu = \mathbbm{1}$.

@@

In our work we adopt as learning strategy the Proximal Policy Optimization (PPO) developed in \textcite{ppo} and \textcite{gae}. PPO is an actor-critic algorithm well-suited for continuous control problems. The term \virgolette{actor-critic} means that it estimates the value function  $J$, which represents the expected value of future rewards, and updates the policy in the direction suggested by $J$. In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$. The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent algorithm in order to maximize an objective function $L^{\text{PPO}}$. The PPO objective is defined as
\begin{equation}
	L^{\text{PPO}} = L^\pi - c_1 L^J + c_2 L^{E}
\end{equation}
where $L^\pi$ is the policy surrogate, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm. We refer to \textcite{ppo} for a deeper description of those terms. 
 We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), where the set of parameters $\theta$ are the hidden weights of the networks. In the following section we describe the the way we have formalized the TVS problem in the Reinforcement Learning framework.

 \subsection{Episode Description: Agent-Environment Interaction}
 We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=0$ and $T_m=T$ maturity of the option. At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^\theta_k$.
 In our case the agent can choose the composition of the risky asset portfolio, so that the policy is the asset allocation weights $\alpha$. From the HJB equation we have derived in section \Cref{subsec:HJB}, we select as state of th environment the following block
 \begin{equation}
 	s_{k} \coloneqq \left[T_k,I_{T_k}, S_{T_k}\right] \quad \forall T_k \in \mathcal{T}.
 \label{eq:state}\end{equation}
 This state contains all the informations needed by the agent to take an optimal action, leading to the maximum option price on the TVS. 
 Once the agent has selected the action $a_k$, it receives the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment defined by
 \begin{equation}
 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)],
 \label{eq:reward}\end{equation}
where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is the BS price as proxy of the residual option price defined by
\begin{equation}
	V_{\text{BS}}(T_k) \coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}}),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \quad \text{and} \quad V(T_0)=0
\end{equation} 
with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} calculated in the state $s_k$. The parameter $\gamma$ plays the role of a discount rate in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future. Choosing $\gamma<1$ introduces a bias in the final result.
If we take the cumulative reward per episode and set $\gamma =1$ we obtain
\begin{equation}
	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
\end{equation} 
which is equal to the intrinsic value of the option. This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we think that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.

@@

\subsection{Policy and Action Parameterization}
As discussed before, PPO parameterizes the agent policy by means of an artificial neural network. In our case we use a feed forward neural network (FFNN), a network with an acyclic topology and completely characterized by the following hyper-parameters: number of hidden layers $N$, number of neurons per hidden layer $M$, activation function per hidden layer and output activation function. The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$. We work with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
\begin{equation}
	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}).
\end{equation}
In the \textsc{Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases. The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.

Another important topic is how to parameterize the agent action in order to satisfy some specific constraints on the risky asset allocation weights. In fact, if on one hand with free $\alpha$ it is sufficient to set bounds in the action space, on the other hand the presence of constraints on the risky strategy imposes the need to parameterize the output of the ANN.
In this work we study the constraint where the agent can invest in the risky market adopting long positions only:
\begin{equation}
	\alpha \cdot \mathds{1} = 1 \quad \alpha^i >0 \quad \forall i=1,\dots,n.
\end{equation} 
We study two different action parameterizations for this kind of constraint.
\begin{itemize}
	\item \textbf{$\bm{(n-1)}$-spherical angles}: we rewrite $\alpha^j = (x^i)^2$ and we use $(n-1)$-spherical coordinates with unitary radius. Thus the strategy is parameterized by a vector of angles $\phi = (\phi^1,\dots,\phi^{n-1})^T$ as follows
	\begin{equation}
		\begin{aligned}
			\alpha^1 &\coloneqq \cos^2 (\phi^1), \\
			\alpha^2 &\coloneqq \sin^2(\phi^1)\cos^2(\phi^2),\\
			&\vdots \\
			\alpha^{n-1} &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\cos^2(\phi^{n-1}),\\
			\alpha^n &\coloneqq \sin^2(\phi^1)\dots\sin^2(\phi^{n-2})\sin^2(\phi^{n-1}).
		\end{aligned}
		\label{eq:euler_angles}\end{equation}
	where $\phi^i\in[0,\pi/2]$ to avoid oscillations in $\alpha$ that could slow down the training. Thus, with this parameterization, the ANN sample a vector of angles ($a = \phi$) and thus the action space is reduced by one dimension.
	\item \textbf{Normalized action}: we restrict the agent actions to a subset of $(\mathbb{R}^+)^n$ and then normalize the action to obtain the corresponding risky allocation strategy through the formula
	\begin{equation}
		\alpha_t = \frac{a_t}{a_t \cdot \mathds{1}}.
		\label{eq:normalized_action}\end{equation}
\end{itemize}

>>>>

\documentclass[11pt]{article}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[outer=3.75cm, marginparwidth=4.5cm]{geometry}
%\usepackage{marginnote}
\usepackage{amssymb,amsmath,amsthm,latexsym,amsfonts,amscd,dsfont,enumerate,bbm,mathabx}
\usepackage{graphicx,tikz}
\usepackage{draftwatermark}
\usepackage{enumerate} 
% \usepackage{listings}
% \usepackage{booktabs}
\usepackage{mathtools}
% \usepackage[gen]{eurosym}
% \usepackage{algorithm}
% \usepackage{algpseudocode}
\usepackage{soul}
\usepackage{comment}
\usepackage{bm}
\usepackage[font=small,labelfont=bf]{caption}
\graphicspath{{Local_Volatility/}{Black_Scholes}}
% \setbeamertemplate{bibliography item}[text]
% \renewcommand*{\bibfont}{\scriptsize}

@@

\author{
	Roberto Daluiso\thanks{Intesa Sanpaolo Milan, roberto.daluiso@intesasanpaolo.com}
	\and 
	Emanuele Nastasi\thanks{Marketz S.p.A., emanuele.nastasi@marketz.eu}
	\and Andrea Pallavicini\thanks{Imperial College London and Intesa Sanpaolo Milan, a.pallavicini@imperial.ac.uk} 
	\and
	Stefano Polo \thanks{Intesa Sanpaolo Milan, stefano.polo@intesasanpaolo.com} \thanks{We wish to thank the Italian computing centre Cineca, which approved our ISCRA C project and provided us the high-computing resources of Marconi100 for the numerical simulations of this work.}% indicates same as 2nd thanks
}

@@

In literature TVSs are tested to investigate their performances in term of realized returns, see for instance \textcite{Hocquard28} and \textcite{PERCHET}, and the soundness of the volatility targeting algorithm, as described in \textcite{Kim}. Moreover, in pricing literature, derivative contracts on TVS, known as target volatility options (TVOs), are studied, see in particular \textcite{Di_Graziano}, \textcite{Grasselli}, and \textcite{Albeverio}.

In this contribution we face a new topic never dealt with in the TVS literature: we deal with the funding costs coming from hedging the risky assets underlying the TVS. We consider the point of view of a bank selling a call option to a portfolio manager as protection on the capital invested in a TVS. The portfolio manager has the freedom of changing the relative weights of the risky assets during the life of the TVS. Since the risky assets have different hedging costs, the bank shall adjust the price of the protection to include them in the worst-case scenario, i.e. the most expensive strategy from the financing costs point of view. Hence, the pricing problem becomes a dynamical control problem over the risky portfolio composition. In our contribution we describe the dynamical control problem, and we derive an analytical solution in the simple case of the risky assets driven by a Black-Scholes (BS) model \citep{BlackScholes}. Then, we tackle the problem in the general case of a local volatility (LV) model \citep{Dupire94pricingwith,LVfirst} by using Reinforcement Learning (RL) algorithms, in particular the Proximal Policy Optimization developed in \textcite{ppo}.
\\

The paper is organized ad follows. In \Cref{sec:TVS} we describe the dynamics of a TVS in presence of valuation adjustments. Then, in \Cref{sec:Derivative} we study the case of derivative contracts on TVSs, we describe the dynamical control problem giving the expression of the Hamilton-Jacobi-Bellman equation and we derive the analytical results for plain vanilla options under BS dynamics. In \Cref{sec:RL} we illustrate how we have applied RL to solve the dynamic control problem, giving a deep description of the algorithm we have built. We conclude the paper with \Cref{sec:Numerical_results} where we present the numerical results obtained in this work for the Black and Scholes model and the local volatility (LV) one.

@@

    \frac{dI_t^{\text{MP}}}{I_t^{\text{MP}}} \coloneqq \left(r_t - \ell_{\alpha} \left(t,I_t^{\text{MP}}\right)\right)dt + \bar{\sigma}dW_t^{\text{MP}},

@@

    \ell_\alpha \left(t,K\right) \coloneqq \bar{\sigma} \mathbb{E}_0\left[\frac{\mu_t \cdot \alpha_t}{\|\alpha_t \cdot \nu_t \|}\bigg|I_t = K\right],

@@

\subsection{Black and Scholes Model}\label{subsec:BS_model}
In the Black and Scholes model with deterministic rates, we can work with empty $X_t$, since in this case the portfolio dynamics \eqref{eq:TVS_last} is Markovian, leading to an optimal strategy $\alpha_t^*$ which depends in principle only on $I_t$. As a consequence, the local drift can be written as
\begin{equation}
     \ell_\alpha \left(t,K\right) = \bar{\sigma}   \frac{\mu\left(t\right) \cdot \alpha\left(t,K\right)}{\|\alpha\left(t,K\right) \cdot \nu\left(t\right)\|},  
\end{equation}
so that the optimization problem can be solved looking only at the Markovian projection without simulating all the Brownian motions $W_t$. Notice that we are indicating the dependency on time in parenthesis to highlight that in this formula all the quantities are deterministic function of time.

@@

\begin{proposition}
When the underlying securities follow a Black and Scholes (BS) model with deterministic rates, the optimal strategy for a monotone payoff consists in minimizing the local drift function, independently of the current state $I_t$
\begin{equation}
    \alpha^*(t) \coloneqq \argmin_\alpha \frac{\alpha \cdot \mu(t)}{\|\alpha \cdot \nu(t) \|}.
\label{eq:BS_optimal_strategy}\end{equation}
\end{proposition}
The absence of stochastic elements in \Cref{eq:BS_optimal_strategy} makes the optimal strategy known \textit{a priori} with no simulation needed; in fact one can solve the optimization problem once for all $\forall t \in \mathcal{T}$ just looking at the market data $\mu(t)$ and $\nu(t)$ for the securities. Once $\alpha^*$ is known, then one can price the payoff by the following BS formula
\begin{equation}
	V_0^{\text{BS}} = BS(F^{\text{TVS}}(0,T;\alpha^*,K), K, T, \bar{\sigma}, D(0,T;\zeta)),
\label{eq:bs_closed_solution}\end{equation}
where $F^{\text{TVS}}(t,T;\alpha,K)$ is the TVS forward curve defined by
\begin{equation}
	F^{\text{TVS}}(t,T;\alpha, K) = I_t \exp\left[\int_t^T\left(r(u)-\ell_\alpha \left(u,K\right)\right)du\right],
\end{equation}
while $BS(F,K,T,\sigma,D)$ is the standard BS formula for an European option with forward curve $F$, strike $K$, time to maturity $T$, volatility $\sigma$ and discount factor $D$.  
\subsubsection{Black's Closed form solution}
In absence of constraints on the allocation strategy, we are able to derive a closed form solution to the BS problem \eqref{eq:BS_optimal_strategy}
\begin{lemma}
	Let be $\mu, \alpha \in \mathbb{R}^n$, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix and $\Sigma\coloneqq \nu \tr{\nu}$. Then the closed solution of the optimization problem \eqref{eq:BS_optimal_strategy} is 
	\begin{equation}
			\alpha^* = - \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
	\end{equation}
\end{lemma}
\begin{proof}
Since the argument of the minimum \eqref{eq:BS_optimal_strategy} is zero-homogeneous, then we can rewrite the problem as
\begin{equation}
		\begin{aligned}
			&\text { minimize } \alpha \cdot \mu\\
			&\text { subject to } \|\alpha \cdot \nu\|^2=1
	\end{aligned}
\end{equation}
By setting the Lagrangian function associated with the problem
\begin{equation}
	\mathcal{L}\left(\alpha, \lambda\right)=\alpha \cdot \mu-\lambda\left(\|\alpha \cdot \nu\|^2-1\right) \, ,
\end{equation}
we obtain the first order conditions
\begin{eqsys}
	\frac{\partial \mathcal{L}}{\partial \alpha}=\mu -2\lambda \Sigma \cdot \alpha=0 \\
	\frac{\partial \mathcal{L}}{\partial \lambda}= \|\alpha \cdot \nu\|^2-1=0
	\, ,
\end{eqsys}
Then, by applying simple algebra, we obtain the analytical form of the free optimal strategy
\begin{equation}
	\alpha^* = \pm \frac{\Sigma^{-1} \cdot \mu}{\|(\Sigma^{-1} \cdot \mu)\cdot \nu\|}
\end{equation}
We take the minus sign to get the minimum value of the TVS local drift. 
\end{proof}
\subsubsection{Active Asset (or Bang Bang) Solution}
A closed form solution to the minimization of the local drift correction \eqref{eq:BS_optimal_strategy} can also be derived in the common case that all costs of carry are nonnegative and the only constraint on portfolio weights is nonnegativity, which would mean a long only strategy by the fund manager.

\begin{lemma}
	Let $\mu \in \mathbb{R}^n$ be a vector with nonnegative components, $\nu \in \mathbb{R}^{n\times n}$ be a full rank matrix, and $\Sigma = \nu \tr{\nu}$. Then
	\begin{equation}\label{eq:inf}
		\inf_{\alpha \in \mathbb{R}_{+}^n \setminus \{0\}} \frac{\alpha \cdot \mu}{\norm{\alpha \cdot \nu}} = \min_{i\leq n} \frac{\mu_i}{\sqrt{\Sigma_{ii}}};
	\end{equation}
	if $\bar{\imath}$ is the index which realizes the min, then the infimimum is realized by a vector concentrated on the $\bar{\imath}$ component: $\alpha_i = \delta_{i\bar{\imath}}$. 
\end{lemma}
\begin{proof}
	Let us first consider the case in which $\mu = \mathbbm{1}$. Since the argument of the infimum is zero-homogeneous, normalizing by $\alpha \cdot \mathbbm{1} > 0$ we can restrict to the affine hyperspace $\{\alpha \cdot \mathbbm{1} = 1\}$, where the minimization \eqref{eq:inf} reduces to the maximization of its denominator: the required infimum will be square root of the reciprocal of
	\[
	\sup\left\{\norm{\alpha \cdot \nu}^2 \mid \alpha \in \mathbb{R}_{+}^n,\, \alpha \cdot \mathbbm{1} = 1\right\}.
	\]
	Now we can note that $\Sigma$ is positive definite, hence $\Sigma_{ij} < \sqrt{\Sigma_{ii}\Sigma_{jj}} \leq \Sigma_{\bar{\imath}\bar{\imath}}$, which implies
	\[
	\norm{\alpha \cdot \nu}^2 = \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{ij} \leq \sum_{i,j=1}^{n} \alpha_i \alpha_j \Sigma_{\bar{\imath}\bar{\imath}} = \Sigma_{\bar{\imath}\bar{\imath}}
	\]
	because $\sum_i \alpha_i = 1$. Since we trivially have equality for $\alpha_i = \delta_{i\bar{\imath}}$, this concludes the proof of the case $\mu = \mathbbm{1}$.

@@

In our work we adopt as learning strategy the proximal policy optimization (PPO) developed in \textcite{ppo} and \textcite{gae}. PPO is an actor-critic algorithm well-suited for continuous control problems. The term \virgolette{actor-critic} means that it estimates the value function  $J$, which represents the expected value of future rewards, and updates the policy in the direction suggested by $J$. In PPO the policy distribution and the value function are parameterized by a set of parameters $\theta$: $J^\theta$, $\pi^\theta$. The PPO collects a small batch of experiences interacting with the environment to update $\theta$ by computing an estimator of the policy gradient and plugging it into a stochastic gradient ascent (SGA) algorithm in order to maximize an objective function $L^{\text{PPO}}$. The PPO objective is defined as
\begin{equation}
	L^{\text{PPO}} = L^\pi - c_1 L^J + c_2 L^{E}
\end{equation}
where $L^\pi$ is the policy surrogate, $L^J$ the value function error term and $L^E$ the entropy bonus, while $c_1$ and $c_2$ are hyper-parameters of the algorithm. We refer to \textcite{ppo} for a deeper description of those terms. 
 We use the PPO implementation found in \textsc{OpenAI Baselines} \citep{baselines}, where the parameterization procedure is performed by using artificial neural networks (ANNs), where the set of parameters $\theta$ are the hidden weights of the networks. In the following section we describe the the way we have formalized the TVS problem in the Reinforcement Learning framework.

 \subsection{Episode Description: Agent-Environment Interaction}
 We consider an episode $\tau$ of length $m+1$ that takes place on a discrete time-grid of fixing times expressed in year fractions $\mathcal{T}:=\{T_0,...,T_k,...,T_m\}$ with $T_0:=0$ and $T_m=T$ maturity of the option. At a given episodic time $T_k$ the RL agent interacts with the environment: it receives a representation of the environment called state $s_k$ and on the basis of that it selects an action $a_k$ sampling from the current policy $\pi^\theta_k$.
 In our case the agent can choose the composition of the risky asset portfolio, so that the policy is the asset allocation weights $\alpha$. From the HJB equation we have derived in section \Cref{subsec:HJB}, we select as state of th environment the following block
 \begin{equation}
 	s_{k} \coloneqq \left[T_k,I_{T_k}, S_{T_k}\right] \quad \forall T_k \in \mathcal{T}.
 \label{eq:state}\end{equation}
 This state contains all the informations needed by the agent to take an optimal action, leading to the maximum option price on the TVS. 
 Once the agent has selected the action $a_k$, it receives the next time $T_{k+1}$ a reward $r_{k+1}$ generated by the environment defined by
 \begin{equation}
 	r_{k+1} \coloneqq  \gamma^{k}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)],
 \label{eq:reward}\end{equation}
where $\gamma \in[0,1]$ is an hyper-parameter of the PPO, while $V_{\text{BS}}(T_k)$ is the BS price as proxy of the residual option price defined by
\begin{equation}
	V_{\text{BS}}(T_k) \coloneqq BS(F^{\text{TVS}}(T_k,T;\alpha^*_{\text{BS}},K),K,T-T_k,\bar{\sigma},D(T_k,T;\zeta)) \quad \text{and} \quad V(T_0)=0
\end{equation} 
with $\alpha^*_{BS}$ the BS optimal strategy \eqref{eq:BS_optimal_strategy} calculated in the state $s_k$. The parameter $\gamma$ plays the role of a discount rate in the sense that, as $\gamma$ approaches to zero, the RL agent will tend to maximize immediate rewards while neglecting possible larger rewards in the future. Choosing $\gamma<1$ introduces a bias in the final result.
If we take the cumulative reward per episode and set $\gamma =1$ we obtain
\begin{equation}
	R(\tau) = \sum_{k=0}^{m-1} r_{k+1} \underset{\gamma =1}{=} \sum_{k=0}^{m-1}[V_{\text{BS}}(T_{k+1})-V_\text{\text{BS}}(T_k)] = V_{BS}(T) = (I_T-K)^+,
\end{equation} 
which is equal to the intrinsic value of the option. This result does not depend on the definition of $V_{\text{BS}}$  $\forall T_k<T$, but we think that the closer $V_{\text{BS}}$ is to the value function, the easier the agent is in learning.

@@

\subsection{Policy and Action Parameterization}
As discussed before, PPO parameterizes the agent policy by means of an artificial neural network. In our case we use a feed forward neural network (FFNN), a network with an acyclic topology and completely characterized by the following hyper-parameters: number of hidden layers $N$, number of neurons per hidden layer $M$, activation function per hidden layer and output activation function. The number of output neurons is fixed by the action space dimension, which is the number of assets $n$ in the risky portfolio, while the number of  input neurons is determined by the dimensionality of the state \eqref{eq:state}, $n+2$. We work with a stochastic policy whose functional form is a multivariate diagonal Gaussian distribution where the mean $\mu^\theta(s_k)$ is the output vector of the FFNN and the log-standard deviation $\log\sigma$ is an external parameter
\begin{equation}
	\pi^\theta(s_k) \sim \mathcal{N}(\mu^\theta(s_k),e^{\log\sigma}).
\end{equation}
In the \textsc{OpenAI Baselines} implementation of PPO, the parameter $\log\sigma$ is state-independent, but it decreases as the number of PPO update iterations increases. The idea is that the log-standard deviation will be higher at the beginning of the training phase in order to guarantee a good exploration of the action space while it will be lower at the ending phase to avoid too much noise in the proximity of the optimal policy.
